### 2020-12-19 paper/localpsf.tex
editing weighting functions section to be infinite dimensional
figured out clear proof for infinite dimensional weighting function theorem

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense. Theorem \ref{thm:smoothest_interpolant} relies on some straightforward prerequisites that we present in Definition \ref{def:stiffness_mass}, Proposition \ref{prop:equivalent_writing_of_weighting_optimization}, and Corollary \ref{cor:linearity_of_solution_operator}. In Theorem \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.

In Section \ref{sec:computation_of_weighting_functions}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. Constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added. However, we present an efficient incremental algorithm for updating these weighting functions when a new point is added, so that we only need to perform two new Poisson PDE solves, rather than $2(q+1)$ new PDE solves.

It is important to note that functions in $H^2$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.

\begin{prop}
	\label{prop:equivalent_writing_of_weighting_optimization}
	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
	\end{aligned}
	\end{equation}
	may be rewritten as follows:
	\begin{equation}
		\label{eq:generic_optimization}
		\begin{aligned}
		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
		\text{such that} &\quad 
		\mathcal{P} u = f.
		\end{aligned}
	\end{equation}
\end{prop}

\begin{thm}
	\label{thm:smoothest_interpolant}
Here let $u$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation}
for an arbitrary but fixed value of $y$, where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
		&= u.
	\end{align*}
 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Corollary \ref{cor:linearity_of_solution_operator}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

@@

@@

	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
	\begin{equation}
	\label{eq:discrete_poisson_optimization_2}
	\begin{aligned}
	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
	\end{aligned}
	\end{equation}
	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
	\begin{equation*}
		\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
	\end{equation*}
	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
	\begin{equation*}
		0 = \nabla \mathcal{L} = 
		\begin{bmatrix}
			\nabla_{v} \mathcal{L} \\
			\nabla_{\lambda} \mathcal{L} \\
			\nabla_\alpha \mathcal{L}
		\end{bmatrix} 
		= \begin{bmatrix}
			\mathcal{M}^{-1} v + \mathcal{K}^+ \mathcal{P}^T \lambda \\
			\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
			\mathbf{1}^T \lambda
		\end{bmatrix}.
	\end{equation*}
	This equation may be rewritten as the following block $3 \times 3$ linear system:
	\begin{equation*}
		\begin{bmatrix}
			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
			\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
			0 & \mathbf{1}^T & 0
		\end{bmatrix}
		\begin{bmatrix}
			v \\ \lambda \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ e \\ 0
		\end{bmatrix}.
	\end{equation*}
	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
	\begin{equation*}
		\begin{bmatrix}
			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
			& -S & \mathbf{1} \\
			0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
		\end{bmatrix}
		\begin{bmatrix}
			v \\ \lambda \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ e \\ \mathbf{1}^T S^{-1} e
		\end{bmatrix}
	\end{equation*}
	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \mathcal{K}^+ \mathcal{P}^T$. From this block triangular system, we read off the solution as:
	\begin{align*}
		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
		\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
	\end{align*}
	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
	\begin{equation*}
		\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	as required.
\end{proof}

@@

			Form the weighting function
			$$w_i = -\sum_{k=1}^r \lambda_k \theta_k + \alpha C,$$
			where we recall that $C$ is the constant function $C(x)=1$ on $\Omega$.
		}
	}
	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
	\label{alg:weighting_functions_incremental}
\end{algorithm2e}

@@

>>>>

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense. In Theorem \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise. These theorems rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. We prove \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}. Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. The Poisson PDE solves can be performed incrementally, so that we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.

It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.

\begin{lem}
	\label{lem:T_is_linear}
	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
\end{lem}

\begin{thm}
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, and let $u$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
		&= u.
	\end{align*}
 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

@@

			Form the weighting function $w_i$ given by
			$$w_i(x) = \alpha - \sum_{k=1}^r \lambda_k \theta_k(x).$$
		}
	}
	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
	\label{alg:weighting_functions_incremental}
\end{algorithm2e}

\begin{thm}
\label{thm:wi_alg_is_correct}
	The weighting functions $w_i$ computed by Algorithm \ref{alg:weighting_functions_incremental} solve optimization problem \ref{eq:wi_optimization_problem}.
\end{thm}

@@

	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
	\begin{equation}
	\label{eq:discrete_poisson_optimization_2}
	\begin{aligned}
	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
	\end{aligned}
	\end{equation}
	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
	\begin{equation*}
		\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
	\end{equation*}
	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
	\begin{equation*}
		0 = \nabla \mathcal{L} = 
		\begin{bmatrix}
			\nabla_{v} \mathcal{L} \\
			\nabla_{\lambda} \mathcal{L} \\
			\nabla_\alpha \mathcal{L}
		\end{bmatrix} 
		= \begin{bmatrix}
			\mathcal{M}^{-1} v + \left(\mathcal{K}^+\right)^T \mathcal{P}^T \lambda \\
			\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
			\mathbf{1}^T \lambda
		\end{bmatrix}.
	\end{equation*}
	This equation may be rewritten as the following block $3 \times 3$ linear system:
	\begin{equation*}
		\begin{bmatrix}
			\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
			\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
			0 & \mathbf{1}^T & 0
		\end{bmatrix}
		\begin{bmatrix}
			v \\ \lambda \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ e \\ 0
		\end{bmatrix}.
	\end{equation*}
	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
	\begin{equation*}
		\begin{bmatrix}
			\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
			& -S & \mathbf{1} \\
			0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
		\end{bmatrix}
		\begin{bmatrix}
			v \\ \lambda \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ e \\ \mathbf{1}^T S^{-1} e
		\end{bmatrix}
	\end{equation*}
	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \left(\mathcal{K}^+\right)^T \mathcal{P}^T$. From this block triangular system, we read off the solution as:
	\begin{align*}
		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
		\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
	\end{align*}
	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
	\begin{equation*}
		\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	as required.
\end{proof}

@@

@@

\begin{prop}
	\label{prop:solution_operator_is_linear}
	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial_old}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
	\end{aligned}
	\end{equation}
	may be rewritten as follows:
	\begin{equation}
	\label{eq:generic_optimization}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
	\text{such that} &\quad 
	\mathcal{P} u = f.
	\end{aligned}
	\end{equation}
\end{prop}
